---
title: "Body size spatial analysis"
author: "Ben Letcher"
date: "8/4/2020"
output:
  html_document:
    code_folding: hide
---

***
### Outline:
1. Copied from ecoDrought_conditionAnalysis project
2. Hierarchical joy plots.
2. Residual means by waterbody - how much variation can we explain with lmer()?

***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE, warning = FALSE, message = FALSE, fig.width = 12, fig.height = 10) #, cache = TRUE)
```

```{r load libraries, echo = FALSE, message = FALSE}
library(tidyverse)
library(lme4)
library(broom)
library(kableExtra)
library(modelr)
library(data.table)
library(rlang)
library(tidyfast)
library(knitr)
```

If newData, then will load raw files, parse, filter and save as 'dat'
```{r newData}
newData = FALSE
date <- 'June4' #suffix for data subdirectory
```


Only need to run this with new data - data output as 'dat'
```{r fetch data}
if(newData) {

  #bring in all of the fish size data
  fish <- list.files(paste0("D:/projects/ecoDrought/bodySize/All_Size_inputs/Size_inputs_", date), pattern = "*.txt", full.names = T)
  
  txt_files_df <- lapply(fish, function(x) {read.table(file = x, header = T, sep =",")})
  
  all_fish <- do.call("rbind", lapply(txt_files_df, as.data.frame))
  
  ##bring in the covariate data
  covs <- read.csv("D:/projects/ecoDrought/bodySize/All_Size_inputs/covariates_size_analysis.csv")
  
  #merge fish data with covariate data
  all_fish_init<-merge(all_fish, covs, by = "RALLID")
  
  #Select only those fish of interest
  
  #species of interest    
  sinc <- c("Bonneville Cutthroat Trout","Colorado River Cutthroat Trout","Redband Trout","Bull Trout","Brook Trout","Mountain Whitefish","Brown Trout","Westslope Cutthroat Trout", "Yellowstone Cutthroat Trout","Rainbow Trout")
  
  #exclude weir data
  nw <- c("Lake","stream")
  
  dat0 <- all_fish_init %>%
    filter(species %in% sinc, site_type %in% nw,
           !is.na(Weight), Length > 75) %>%
    rename(weight = Weight, length = Length) %>%
    mutate(log10Weight = log10(weight), log10Length = log10(length))
  
  # Species-specific LWE regressions so we can add residuals to dataset
  lwModel <- function(df) {
    lm(log10Weight ~ log10Length, data = df)
  }
  
  datBySpecies <- dat0 %>%
    group_by(species) %>% 
    nest() %>%
    mutate(model = map(data, lwModel))
           
  datBySpecies <- datBySpecies %>% mutate(resids = map2(data, model, add_residuals))

  dat <- dt_unnest(datBySpecies, resids) # unnest() from tidyverse is so slow it never finishes
  
  dat <- dat %>%
    group_by(location) %>%
    mutate(daStd = (log(drain_area_km2) - (mean(log(drain_area_km2)))) / sd(log(drain_area_km2)),
           sFStd = (log(summerF) - (mean(log(summerF + 0.0001), na.rm=TRUE))) / sd(log(summerF + 0.0001), na.rm=TRUE),
           sTStd = (log(summerT) - (mean(log(summerT), na.rm=TRUE))) / sd(log(summerT), na.rm=TRUE),
           slopeStd = (SLOPE - mean(SLOPE)) / sd(SLOPE),
           rdStd = (log(rel_density) - (mean(log(rel_density), na.rm=TRUE))) / sd(log(rel_density), na.rm=TRUE)) %>%
    ungroup()
  
  dat$datePosix <- as.POSIXlt(strptime(dat$date, "%m/%d/%Y %H:%M:%S"))
  dat$doy <- lubridate::yday(dat$datePosix)
  
  speciesList <- unique(dat$species)
  
  save(dat, sinc, file =paste0("D:/projects/ecoDrought/bodySize/All_Size_inputs/Size_inputs_", date,"/dat.R"))
  
  rm(all_fish_init, all_fish, dat0)
}
```

```{r load data if no new data}
if(!newData){
  load(paste0("D:/projects/ecoDrought/bodySize/All_Size_inputs/Size_inputs_", date,"/dat.R"))
  speciesList <- unique(dat$species)
}
```

***

## 1. Spatial distribution of LW residuals by species
Spatially-nested lmer() to explore spatial structuring of variation in body sizes
Location nested within HUC12 nested within HUC8 nested within species

***

```{r spatially-nested lmer by species}



mod0 <- lmer(log10Length ~ 1 + (1|Location/HUC12_NAME/HUC8_Name/species), data = dat)




# build up hierarchical random effects for each species 
## Hierarchical graphs for huc12/huc8/location 
  lmerModel <- function(df) {
    lmer(resid ~ 1 + (1|location/HUC8_Name/HUC12_NAME), data = df)
  }

  reBySpecies <- function(mod, species){
    re <- as.data.frame(ranef(mod))

    re12 <- re %>%
      filter(grpvar == 'HUC12_NAME:(HUC8_Name:location)') %>%
      mutate(huc12 = str_split(grp, ":", simplify = TRUE)[,1],
             huc8 = str_split(grp, ":", simplify = TRUE)[,2],
             loc = str_split(grp, ":", simplify = TRUE)[,3]) %>%
      rename(condval12 = condval)
  
    re8 <- re %>%
      filter(grpvar == 'HUC8_Name:location') %>%
      mutate(loc = str_split(grp, ":", simplify = TRUE)[,2],
             huc8 = str_split(grp, ":", simplify = TRUE)[,1]) %>%
      rename(condval8 = condval)
    
    reLoc <- re %>%
      filter(grpvar == 'location') %>%
      mutate(loc = grp) %>%
      rename(condvalLoc = condval)%>% 
      select(condvalLoc, loc)
    
    reAll1 <- left_join(re8, reLoc) %>% select(condval8, loc, huc8, condvalLoc)
    reAll <- left_join(re12, reAll1) %>% arrange(loc, huc8, huc12)
    reAll <- reAll %>%
      mutate(condvalLoc8 = condvalLoc + condval8,
             condvalLoc812 = condvalLoc + condval8 + condval12)
    
    return(reAll)
  }
  
  hierRECharts <- function(d, species) {
      gg <- ggplot(d, aes(condvalLoc812, huc12)) +
      geom_point(color = '#ff9a36') +
      geom_point(aes(condvalLoc8, huc12), color = '#3f7bc4') +
      geom_point(aes(condvalLoc, huc12, shape = factor(loc)), color = '#bdb2a8') +
      geom_vline(xintercept = 0) +
      scale_y_discrete(guide = guide_none(title = NULL)) +
      scale_x_continuous("Residual from species-specific LW relationship") +
      ggtitle(species) +
      facet_wrap(~huc8)
    
    ggsave(paste0("D:/projects/ecoDrought/bodySize/All_Size_inputs/ecoDrought_conditionAnalysis/charts/hierResids_",species,".png"))
    return(gg)
  }
  
##  
    # Charts of nested random effects 
  datBySpeciesRE <- dat %>%
    filter(species %in% speciesList[c(1,3,4,5,7)]) %>%
    #filter(species %in% speciesList[5]) %>% #
    group_by(species) %>% 
    nest() %>%
    mutate(model = map(data, lmerModel)) %>%
    mutate(RE = map2(model, species, reBySpecies)) %>%
    mutate(charts = map2(RE, species, hierRECharts))
    
  datBySpeciesRE$charts
  
  ## SDs of nested levels ################
  datBySpeciesRESD <- datBySpeciesRE %>%
    mutate(tidy = map(model, broom::tidy)) %>% 
    dt_unnest(tidy) %>%
    select(species, estimate, term, group) %>%
    filter(!term == '(Intercept)')
  
  ggSD <- ggplot(datBySpeciesRESD, aes(species,estimate, fill = group)) +
    geom_col(position = "dodge") +
    scale_y_continuous("Standard deviation of nested LW residuals")
  ggSD
  
  ## plot RE across species
  datBySpeciesRE_RE <- datBySpeciesRE %>%
    dt_unnest(RE)

  # by loc
  ggplot(datBySpeciesRE_RE, aes(condvalLoc812, species)) +
    geom_point(color = '#ff9a36', alpha = 0.2, position = position_nudge(y = 0.2)) +
    geom_point(aes(condvalLoc8, species), color = '#3f7bc4', alpha = 0.2, position = position_nudge(y = -0.2)) +
    geom_point(aes(condvalLoc, species), color = 'black') +
    geom_vline(xintercept = 0) +
    scale_x_continuous("Residual from species-specific LW relationship") +
    facet_wrap(~loc)

  # by huc8
  nByhuc8 <- datBySpeciesRE_RE %>%
    group_by(huc8) %>%
    summarise(nSpp = length(unique(species)))
  
  datBySpeciesRE_RE_SppN <- left_join(datBySpeciesRE_RE, nByhuc8)
  
  ggplot(datBySpeciesRE_RE_SppN %>% filter(nSpp >1), aes(condvalLoc812, species)) +
    geom_point(color = '#ff9a36', alpha = 0.2, position = position_nudge(y = 0.2)) +
    geom_point(aes(condvalLoc8, species), color = '#3f7bc4', position = position_nudge(y = -0.2)) +
    geom_point(aes(condvalLoc, species), color = 'black') +
    geom_vline(xintercept = 0) +
    scale_x_continuous("Residual from species-specific LW relationship") +
    facet_wrap(loc ~ huc8, labeller = labeller(.multi_line = FALSE))
  
##############################################################################  
  # some species have data for one location only #2 6 8 9
  # Colorado River Cutthroat Trout model has singularity - ignoring for now
  ## Hierarchical graphs for huc12/huc8
  
  lmerModelNoLocation <- function(df) {
    lmer(resid ~ 1 + (1|HUC8_Name/HUC12_NAME), data = df)
  }
  
  # build up hierarchical random effects for each species 
  reBySpecies812 <- function(mod, species){
    re <- as.data.frame(ranef(mod))

    re12 <- re %>%
      filter(grpvar == 'HUC12_NAME:HUC8_Name') %>%
      mutate(huc12 = str_split(grp, ":", simplify = TRUE)[,1],
             huc8 = str_split(grp, ":", simplify = TRUE)[,2]
             ) %>%
      rename(condval12 = condval)

    re8 <- re %>%
      filter(grpvar == 'HUC8_Name') %>%
      mutate(huc8 = grp) %>%
      rename(condval8 = condval) %>% 
      select(condval8, huc8)
    
    reAll <- left_join(re12, re8) %>% arrange(huc8, huc12)
    reAll <- reAll %>%
      mutate(condval8 = condval8,
             condval812 = condval8 + condval12)
    
    return(reAll)
  }
  
  hierREChartsNoLocation <- function(d, species) {
    gg <- ggplot(d, aes(condval812, huc12)) +
      geom_point(color = '#ff9a36') +
      geom_point(aes(condval8, huc12), color = '#3f7bc4') +
      geom_vline(xintercept = 0) +
      scale_y_discrete(guide = guide_none(title = NULL)) +
      scale_x_continuous("Residual from species-specific LW relationship") +
      ggtitle(species) +
      facet_wrap(~huc8)
    
    ggsave(paste0("D:/projects/ecoDrought/bodySize/All_Size_inputs/ecoDrought_conditionAnalysis/charts/hierResids_",species,".png"))
    
    return(gg)
  }  
  
  # Charts of nested random effects 
  datBySpeciesRENoLocation <- dat %>%
    filter(species %in% speciesList[c(2,8,9)]) %>%
    group_by(species) %>% 
    nest() %>%
    mutate(model = map(data, lmerModelNoLocation)) %>%
    mutate(RE = map2(model, species, reBySpecies812)) %>%
    mutate(charts = map2(RE, species, hierREChartsNoLocation))
    
  datBySpeciesRENoLocation$charts 
  
  ## SDs of nested levels ################
  datBySpeciesRENoLocationSD <- datBySpeciesRENoLocation %>%
    mutate(tidy = map(model, broom::tidy)) %>% 
    dt_unnest(tidy) %>%
    select(species, estimate, term, group) %>%
    filter(!term == '(Intercept)')
  
  ggNoLocationSD <- ggplot(datBySpeciesRENoLocationSD, aes(species,estimate, fill = group)) +
    geom_col(position = "dodge") +
    scale_y_continuous("Standard deviation of nested LW residuals") 
  ggNoLocationSD 
  
  ## plot RE across species
  datBySpeciesRENoLocation_RE <- datBySpeciesRENoLocation %>%
    dt_unnest(RE)

  # by huc8
  nNoLocationByhuc8 <- datBySpeciesRENoLocation_RE %>%
    group_by(huc8) %>%
    summarise(nSpp = length(unique(species)))
  
  datBySpeciesRENoLocation_RE_SppN <- left_join(datBySpeciesRENoLocation_RE, nNoLocationByhuc8)
  
  ggplot(datBySpeciesRENoLocation_RE_SppN %>% filter(nSpp >1), aes(condval812, species)) +
    geom_point(color = '#ff9a36', alpha = 0.2, position = position_nudge(y = 0.2)) +
    geom_point(aes(condval8, species), color = '#3f7bc4', position = position_nudge(y = -0.2)) +
    geom_vline(xintercept = 0) +
    scale_x_continuous("Residual from species-specific LW relationship") +
    facet_wrap( ~ huc8)

```


***

## 5. Residual means by waterbody
Analyze means of residuals by waterbody. Speeds up lmer() and reduces df inflation from individual observations.
Remove waterbodies with < 10 observations.
Look in depth at lm() for Brook Trout

***

```{r means by waterbody}

datByWB <- dat %>%
  group_by(species, Lake, WATERBODY, location, HUC8_Name, HUC12_NAME, rdStd, rel_density,
           sFStd, sTStd, daStd, Str_Order, KM_4th, KM_Lake) %>%
  summarize(meanResid = mean(resid, na.rm = TRUE),
            sdResid = sd(resid, na.rm = TRUE),
            n = n(),
            meanLength = mean(log10Length, na.rm = TRUE),
            meanWeight = mean(log10Weight, na.rm = TRUE)) %>%
  filter(n > 10) %>%
  ungroup()

## bkt deep dive
datBKT <- datByWB %>% filter(species == "Rainbow Trout")

b0 <- lmer(meanResid ~ 1 + (1|location/HUC8_Name), data = datBKT %>% filter(Lake == 0))
b1 <- lmer(meanResid ~ 1 + rdStd + (1|location/HUC8_Name), data = datBKT %>% filter(Lake == 0))

AIC(b0, b1) %>% arrange(AIC)
tidy(b1)

ggplot(datBKT %>% filter(Lake == 0), aes(rdStd, meanResid)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = 'lm') +
  facet_wrap(~location)

# summerF, summerT, drain_area_km2, Str_Order, KM_4th,KM_Lake
ggplot(datBKT %>% filter(Lake == 0, location == 'MT'), aes(rdStd, meanResid)) +
  geom_point(aes(color = daStd)) +
  geom_smooth(method = 'lm') +
  scale_color_viridis_c() +
  facet_wrap(~location)

b2 <- lmer(meanResid ~ 1 + daStd + (1|location/HUC8_Name), data = datBKT %>% filter(Lake == 0))
b3 <- lmer(meanResid ~ 1 + daStd * rdStd + (1|location/HUC8_Name), data = datBKT %>% filter(Lake == 0))

AIC(b0, b1, b2, b3) %>% arrange(AIC)
glance(b3)

b4 <- lmer(meanResid ~ 1 + daStd * meanLength + (1|location/HUC8_Name), data = datBKT %>% filter(Lake == 0))
AIC(b0, b1, b2, b3, b4) %>% arrange(AIC)

# are the random effects sucking up all the variance? Try just lm() to see how var can get explained
blm0 <- lm(meanResid ~ 1, data = datBKT %>% filter(Lake == 0))
blm1 <- lm(meanResid ~ 1 + location, data = datBKT %>% filter(Lake == 0))
blm2 <- lm(meanResid ~ 1 + rdStd * location, data = datBKT %>% filter(Lake == 0))
blm3 <- lm(meanResid ~ 1 + daStd * location, data = datBKT %>% filter(Lake == 0))
blm4 <- lm(meanResid ~ 1 + sFStd * location, data = datBKT %>% filter(Lake == 0))
blm5 <- lm(meanResid ~ 1 + sTStd * location, data = datBKT %>% filter(Lake == 0))
blm6 <- lm(meanResid ~ 1 + meanLength * location, data = datBKT %>% filter(Lake == 0))
blm7 <- lm(meanResid ~ 1 + Str_Order * location, data = datBKT %>% filter(Lake == 0))
blm8 <- lm(meanResid ~ 1 + KM_4th * location, data = datBKT %>% filter(Lake == 0))
blm9 <- lm(meanResid ~ 1 + rdStd * daStd * location, data = datBKT %>% filter(Lake == 0))
blm10 <- lm(meanResid ~ 1 + sFStd * sTStd * location, data = datBKT %>% filter(Lake == 0))
blm11 <- lm(meanResid ~ 1 + meanLength * sFStd * location, data = datBKT %>% filter(Lake == 0))
blm12 <- lm(meanResid ~ 1 + meanLength * sTStd * location, data = datBKT %>% filter(Lake == 0))
blm13 <- lm(meanResid ~ 1 + meanLength * sFStd * sTStd * location, data = datBKT %>% filter(Lake == 0))
blmMega <- lm(meanResid ~ 1 + meanLength * sFStd * sTStd * Str_Order * KM_4th * location * sFStd * sTStd, data = datBKT %>% filter(Lake == 0))

AIC(blm0,blm1,blm2,blm3,blm4,blm5,blm6,blm7,blm8,blm9,blm10,blm11,blm12,blm13,blmMega) %>% arrange(AIC)
glance(blm10)
data.frame(tidy(blm10))
glance(blm13)
data.frame(tidy(blm13))
glance(blmMega)

ggplot(datBKT %>% filter(Lake == 0), aes(sFStd, meanResid)) +
  geom_point(aes(color = sTStd)) +
  geom_smooth(method = 'lm') +
  scale_color_viridis_c() +
  facet_wrap(~location)

ggplot(datBKT %>% filter(Lake == 0, location == 'MT'), aes(sFStd, sTStd, z = meanResid)) +
  geom_density2d() 


```

