---
title: "Body size spatial analysis - Hierarchical joy plots"
author: "Ben Letcher"
date: "8/13/2020"
output:
  html_document:
    code_folding: hide
---

***

***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE, warning = FALSE, message = FALSE, fig.width = 12, fig.height = 10) #, cache = TRUE)
```

```{r load libraries, echo = FALSE, message = FALSE}
library(tidyverse)
library(lme4)
library(broom)
library(kableExtra)
library(modelr)
library(data.table)
library(rlang)
library(tidyfast)
library(knitr)
library(ggridges)
```

If newData, then will load raw files, parse, filter and save as 'dat'
```{r newData}
newData = FALSE
date <- 'June4' #suffix for data subdirectory
```


Only need to run this with new data - data output as 'dat'
```{r fetch data}
if(newData) {

  #bring in all of the fish size data
  fish <- list.files(paste0("D:/projects/ecoDrought/bodySize/All_Size_inputs/Size_inputs_", date), pattern = "*.txt", full.names = T)
  
  txt_files_df <- lapply(fish, function(x) {read.table(file = x, header = T, sep =",")})
  
  all_fish <- do.call("rbind", lapply(txt_files_df, as.data.frame))
  
  ##bring in the covariate data
  covs <- read.csv("D:/projects/ecoDrought/bodySize/All_Size_inputs/covariates_size_analysis.csv")
  
  #merge fish data with covariate data
  all_fish_init<-merge(all_fish, covs, by = "RALLID")
  
  #Select only those fish of interest
  
  #species of interest    
  sinc <- c("Bonneville Cutthroat Trout","Colorado River Cutthroat Trout","Redband Trout","Bull Trout","Brook Trout","Mountain Whitefish","Brown Trout","Westslope Cutthroat Trout", "Yellowstone Cutthroat Trout","Rainbow Trout")
  
  #exclude weir data
  nw <- c("Lake","stream")
  
  dat0 <- all_fish_init %>%
    filter(species %in% sinc, site_type %in% nw,
           !is.na(Weight), Length > 75) %>%
    rename(weight = Weight, length = Length) %>%
    mutate(log10Weight = log10(weight), log10Length = log10(length))
  
  # Species-specific LWE regressions so we can add residuals to dataset
  lwModel <- function(df) {
    lm(log10Weight ~ log10Length, data = df)
  }
  
  datBySpecies <- dat0 %>%
    group_by(species) %>% 
    nest() %>%
    mutate(model = map(data, lwModel))
           
  datBySpecies <- datBySpecies %>% mutate(resids = map2(data, model, add_residuals))

  dat <- dt_unnest(datBySpecies, resids) # unnest() from tidyverse is so slow it never finishes
  
  dat <- dat %>%
    group_by(location) %>%
    mutate(daStd = (log(drain_area_km2) - (mean(log(drain_area_km2)))) / sd(log(drain_area_km2)),
           sFStd = (log(summerF) - (mean(log(summerF + 0.0001), na.rm=TRUE))) / sd(log(summerF + 0.0001), na.rm=TRUE),
           sTStd = (log(summerT) - (mean(log(summerT), na.rm=TRUE))) / sd(log(summerT), na.rm=TRUE),
           slopeStd = (SLOPE - mean(SLOPE)) / sd(SLOPE),
           rdStd = (log(rel_density) - (mean(log(rel_density), na.rm=TRUE))) / sd(log(rel_density), na.rm=TRUE)) %>%
    ungroup()
  
  dat$datePosix <- as.POSIXlt(strptime(dat$date, "%m/%d/%Y %H:%M:%S"))
  dat$doy <- lubridate::yday(dat$datePosix)
  
  speciesList <- unique(dat$species)
  
  save(dat, sinc, file =paste0("D:/projects/ecoDrought/bodySize/All_Size_inputs/Size_inputs_", date,"/dat.R"))
  
  rm(all_fish_init, all_fish, dat0)
}
```

```{r load data if no new data}
if(!newData){
  load(paste0("D:/projects/ecoDrought/bodySize/All_Size_inputs/Size_inputs_", date,"/dat.R"))
  speciesList <- unique(dat$species)
}
```

```{r new variables}

dat <- dat %>%
  mutate(strOrderLake = ifelse(Lake == 1, 'Lake', Str_Order))

```

```{r theme_publication}

theme_publication <- function(base_size=14, base_family="") {
  library(grid)
  library(ggthemes)
  (theme_foundation(base_size=base_size, base_family=base_family)
    + theme(plot.title = element_text(face = "bold",
                                      size = rel(1.2), hjust = 0.5),
            text = element_text(),
            panel.background = element_rect(colour = NA),
            plot.background = element_rect(colour = NA),
            axis.title = element_text(face = "bold",size = rel(1)),
            axis.title.y = element_text(angle=90,vjust =2),
            axis.title.x = element_text(vjust = -0.2),
            axis.text = element_text(),
            axis.line = element_line(colour="black"),
            axis.ticks = element_line(),
            panel.grid.major = element_line(colour="#f0f0f0"),
            panel.grid.minor = element_blank(),
            legend.key = element_rect(colour = NA),
            legend.position = "bottom",
            legend.direction = "horizontal",
            legend.key.size= unit(0.2, "cm"),
            #    legend.margin = unit(0, "cm"),
            legend.title = element_text(face="italic"),
            plot.margin=unit(c(10,5,5,5),"mm"),
            strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
            strip.text = element_text(face="bold"),
            panel.border = element_rect(colour = "black")
    ))

}
```


***
## All fish by species
***
```{r joy plots}
ggplot(dat %>% filter(length < 700), aes(x = length, y = species)
       ) +
  geom_density_ridges(
    rel_min_height = 0.01, # cuts off trailing lines
    scale = 1.75 # Amount of overlap among curves
  ) +
  theme_publication()

# with tail probabilities, https://wilkelab.org/ggridges/articles/introduction.html
ggplot(dat %>% filter(length < 700), aes(x = length, y = species, 
                                         fill = 0.5 - abs(0.5 - stat(ecdf))
                                         )
       ) +
  stat_density_ridges(
    rel_min_height = 0.01, # cuts off trailing lines
    scale = 1.75, # Amount of overlap among curves
    geom = "density_ridges_gradient", 
    calc_ecdf = TRUE,
    #### simulate a rug #### Takes a long time and doesn't look good
    # jittered_points = TRUE,
    # position = position_points_jitter(width = 0.05, height = 0),
    # point_shape = '|', point_size = 3, point_alpha = 1, alpha = 0.7
    ########################
  ) +
  scale_fill_viridis_c(name = "Tail probability", direction = -1) +
  theme_publication() +
  theme(
    legend.position = "right",
    legend.direction = "vertical",
    #legend.key.size= unit(0.2, "cm"),
    # legend.margin = unit(0, "cm"),
    legend.title = element_text(face="bold"),
    axis.text.y = element_text(vjust = 0)
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  coord_cartesian(clip = "off") 
```

***
## All fish by species and strOrderLake
***
```{r}


# by strOrderLake
ggplot(dat %>% filter(length < 700, !is.na(strOrderLake)), 
         aes(x = length, y = factor(strOrderLake), 
            fill = 0.5 - abs(0.5 - stat(ecdf))
            )
       ) +
  stat_density_ridges(
    rel_min_height = 0.01, # cuts off trailing lines
    scale = 1.75, # Amount of overlap among curves
    geom = "density_ridges_gradient", 
    calc_ecdf = TRUE,
    #### simulate a rug #### Takes a long time and doesn't look good
    # jittered_points = TRUE,
    # position = position_points_jitter(width = 0.05, height = 0),
    # point_shape = '|', point_size = 3, point_alpha = 1, alpha = 0.7
    ########################
  ) +
  scale_fill_viridis_c(name = "Tail probability", direction = -1) +
  theme_publication() +
  theme(
    legend.position = "right",
    legend.direction = "vertical",
    #legend.key.size= unit(0.2, "cm"),
    # legend.margin = unit(0, "cm"),
    legend.title = element_text(face="bold"),
    axis.text.y = element_text(vjust = 0)
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  facet_wrap(~species)
```

***
## All fish by species, strOrderLake, and Location
***
```{r}

# by strOrderLake and HUC8 and species
ggplot(dat %>% filter(length < 700, 
                      !is.na(strOrderLake), 
                      #species == 'Brook Trout'
                      ), 
         aes(x = length, y = factor(strOrderLake), 
            fill = 0.5 - abs(0.5 - stat(ecdf))
            )
       ) +
  stat_density_ridges(
    rel_min_height = 0.01, # cuts off trailing lines
    scale = 1.75, # Amount of overlap among curves
    geom = "density_ridges_gradient", 
    calc_ecdf = TRUE,
    #### simulate a rug #### Takes a long time and doesn't look good
    # jittered_points = TRUE,
    # position = position_points_jitter(width = 0.05, height = 0),
    # point_shape = '|', point_size = 3, point_alpha = 1, alpha = 0.7
    ########################
  ) +
  scale_fill_viridis_c(name = "Tail probability", direction = -1) +
  theme_publication() +
  theme(
    legend.position = "right",
    legend.direction = "vertical",
    #legend.key.size= unit(0.2, "cm"),
    # legend.margin = unit(0, "cm"),
    legend.title = element_text(face="bold"),
    axis.text.y = element_text(vjust = 0)
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  #facet_wrap(~ HUC8_Name)
  facet_grid(Location ~ species)
```

***
## Brook Trout by strOrderLake, and Location
***
```{r}
# by strOrderLake and HUC8 for bkt
ggplot(dat %>% filter(length < 700, 
                      !is.na(strOrderLake), 
                      species == 'Brook Trout'
                      ), 
         aes(x = length, y = factor(strOrderLake), 
            fill = 0.5 - abs(0.5 - stat(ecdf))
            )
       ) +
  stat_density_ridges(
    rel_min_height = 0.01, # cuts off trailing lines
    scale = 1.75, # Amount of overlap among curves
    geom = "density_ridges_gradient", 
    calc_ecdf = TRUE,
    #### simulate a rug #### Takes a long time and doesn't look good
    # jittered_points = TRUE,
    # position = position_points_jitter(width = 0.05, height = 0),
    # point_shape = '|', point_size = 3, point_alpha = 1, alpha = 0.7
    ########################
  ) +
  scale_fill_viridis_c(name = "Tail probability", direction = -1) +
  theme_publication() +
  theme(
    legend.position = "right",
    legend.direction = "vertical",
    #legend.key.size= unit(0.2, "cm"),
    # legend.margin = unit(0, "cm"),
    legend.title = element_text(face="bold"),
    axis.text.y = element_text(vjust = 0)
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  facet_wrap(~ Location)
  #facet_grid(Location ~ species)

```

***
## Brown Trout by strOrderLake, and Location
***
```{r}
# by strOrderLake and HUC8 for bkt
ggplot(dat %>% filter(length < 700, 
                      !is.na(strOrderLake), 
                      species == 'Brown Trout'
                      ), 
         aes(x = length, y = factor(strOrderLake), 
            fill = 0.5 - abs(0.5 - stat(ecdf))
            )
       ) +
  stat_density_ridges(
    rel_min_height = 0.01, # cuts off trailing lines
    scale = 1.75, # Amount of overlap among curves
    geom = "density_ridges_gradient", 
    calc_ecdf = TRUE,
    #### simulate a rug #### Takes a long time and doesn't look good
    # jittered_points = TRUE,
    # position = position_points_jitter(width = 0.05, height = 0),
    # point_shape = '|', point_size = 3, point_alpha = 1, alpha = 0.7
    ########################
  ) +
  scale_fill_viridis_c(name = "Tail probability", direction = -1) +
  theme_publication() +
  theme(
    legend.position = "right",
    legend.direction = "vertical",
    #legend.key.size= unit(0.2, "cm"),
    # legend.margin = unit(0, "cm"),
    legend.title = element_text(face="bold"),
    axis.text.y = element_text(vjust = 0)
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  facet_wrap(~ Location)
  #facet_grid(Location ~ species)

```

***
## Rainbow Trout by strOrderLake, and Location
***
```{r}
# by strOrderLake and HUC8 for bkt
ggplot(dat %>% filter(length < 700, 
                      !is.na(strOrderLake), 
                      species == 'Rainbow Trout'
                      ), 
         aes(x = length, y = factor(strOrderLake), 
            fill = 0.5 - abs(0.5 - stat(ecdf))
            )
       ) +
  stat_density_ridges(
    rel_min_height = 0.01, # cuts off trailing lines
    scale = 1.75, # Amount of overlap among curves
    geom = "density_ridges_gradient", 
    calc_ecdf = TRUE,
    #### simulate a rug #### Takes a long time and doesn't look good
    # jittered_points = TRUE,
    # position = position_points_jitter(width = 0.05, height = 0),
    # point_shape = '|', point_size = 3, point_alpha = 1, alpha = 0.7
    ########################
  ) +
  scale_fill_viridis_c(name = "Tail probability", direction = -1) +
  theme_publication() +
  theme(
    legend.position = "right",
    legend.direction = "vertical",
    #legend.key.size= unit(0.2, "cm"),
    # legend.margin = unit(0, "cm"),
    legend.title = element_text(face="bold"),
    axis.text.y = element_text(vjust = 0)
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  facet_wrap(~ Location)
  #facet_grid(Location ~ species)

```

# Location by species
```{r}

ggplot(dat %>% filter(length < 700, 
                      !is.na(strOrderLake), 
                      #species == 'Brook Trout'
                      ), 
         aes(x = length, y = Location, 
            fill = 0.5 - abs(0.5 - stat(ecdf))
            )
       ) +
  stat_density_ridges(
    rel_min_height = 0.01, # cuts off trailing lines
    scale = 1.75, # Amount of overlap among curves
    geom = "density_ridges_gradient", 
    calc_ecdf = TRUE,
    #### simulate a rug #### Takes a long time and doesn't look good
    # jittered_points = TRUE,
    # position = position_points_jitter(width = 0.05, height = 0),
    # point_shape = '|', point_size = 3, point_alpha = 1, alpha = 0.7
    ########################
  ) +
  scale_fill_viridis_c(name = "Tail probability", direction = -1) +
  theme_publication() +
  theme(
    legend.position = "right",
    legend.direction = "vertical",
    #legend.key.size= unit(0.2, "cm"),
    # legend.margin = unit(0, "cm"),
    legend.title = element_text(face="bold"),
    axis.text.y = element_text(vjust = 0)
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  facet_wrap(~ species)
  
```

# Huc12 by Location for bkt - doesn't really work
```{r}

ggplot(dat %>% filter(length < 700, 
                      !is.na(strOrderLake), 
                      species == 'Brook Trout'
                      ), 
         aes(x = length, y = HUC8_Name, 
            fill = 0.5 - abs(0.5 - stat(ecdf))
            )
       ) +
  stat_density_ridges(
    rel_min_height = 0.01, # cuts off trailing lines
    scale = 1.75, # Amount of overlap among curves
    geom = "density_ridges_gradient", 
    calc_ecdf = TRUE,
    #### simulate a rug #### Takes a long time and doesn't look good
    # jittered_points = TRUE,
    # position = position_points_jitter(width = 0.05, height = 0),
    # point_shape = '|', point_size = 3, point_alpha = 1, alpha = 0.7
    ########################
  ) +
  scale_fill_viridis_c(name = "Tail probability", direction = -1) +
  theme_publication() +
  theme(
    legend.position = "right",
    legend.direction = "vertical",
    #legend.key.size= unit(0.2, "cm"),
    # legend.margin = unit(0, "cm"),
    legend.title = element_text(face="bold"),
    axis.text.y = element_text(vjust = 0)
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  facet_wrap(~ Location)
  #facet_grid(Location ~ species)


```