---
title: "Body size spatial analysis"
author: "Ben Letcher"
date: "8/4/2020"
output:
  html_document:
    code_folding: hide
---

***
### Outline:
1. Copied from ecoDrought_conditionAnalysis project
2. Hierarchical joy plots.
2. Residual means by waterbody - how much variation can we explain with lmer()?

***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE, warning = FALSE, message = FALSE, fig.width = 12, fig.height = 10) #, cache = TRUE)
```

```{r load libraries, echo = FALSE, message = FALSE}
library(tidyverse)
library(lme4)
library(broom)
library(kableExtra)
library(modelr)
library(data.table)
library(rlang)
library(tidyfast)
library(knitr)
library(ggridges)
```

If newData, then will load raw files, parse, filter and save as 'dat'
```{r newData}
newData = FALSE
date <- 'June4' #suffix for data subdirectory
```


Only need to run this with new data - data output as 'dat'
```{r fetch data}
if(newData) {

  #bring in all of the fish size data
  fish <- list.files(paste0("D:/projects/ecoDrought/bodySize/All_Size_inputs/Size_inputs_", date), pattern = "*.txt", full.names = T)
  
  txt_files_df <- lapply(fish, function(x) {read.table(file = x, header = T, sep =",")})
  
  all_fish <- do.call("rbind", lapply(txt_files_df, as.data.frame))
  
  ##bring in the covariate data
  covs <- read.csv("D:/projects/ecoDrought/bodySize/All_Size_inputs/covariates_size_analysis.csv")
  
  #merge fish data with covariate data
  all_fish_init<-merge(all_fish, covs, by = "RALLID")
  
  #Select only those fish of interest
  
  #species of interest    
  sinc <- c("Bonneville Cutthroat Trout","Colorado River Cutthroat Trout","Redband Trout","Bull Trout","Brook Trout","Mountain Whitefish","Brown Trout","Westslope Cutthroat Trout", "Yellowstone Cutthroat Trout","Rainbow Trout")
  
  #exclude weir data
  nw <- c("Lake","stream")
  
  dat0 <- all_fish_init %>%
    filter(species %in% sinc, site_type %in% nw,
           !is.na(Weight), Length > 75) %>%
    rename(weight = Weight, length = Length) %>%
    mutate(log10Weight = log10(weight), log10Length = log10(length))
  
  # Species-specific LWE regressions so we can add residuals to dataset
  lwModel <- function(df) {
    lm(log10Weight ~ log10Length, data = df)
  }
  
  datBySpecies <- dat0 %>%
    group_by(species) %>% 
    nest() %>%
    mutate(model = map(data, lwModel))
           
  datBySpecies <- datBySpecies %>% mutate(resids = map2(data, model, add_residuals))

  dat <- dt_unnest(datBySpecies, resids) # unnest() from tidyverse is so slow it never finishes
  
  dat <- dat %>%
    group_by(location) %>%
    mutate(daStd = (log(drain_area_km2) - (mean(log(drain_area_km2)))) / sd(log(drain_area_km2)),
           sFStd = (log(summerF) - (mean(log(summerF + 0.0001), na.rm=TRUE))) / sd(log(summerF + 0.0001), na.rm=TRUE),
           sTStd = (log(summerT) - (mean(log(summerT), na.rm=TRUE))) / sd(log(summerT), na.rm=TRUE),
           slopeStd = (SLOPE - mean(SLOPE)) / sd(SLOPE),
           rdStd = (log(rel_density) - (mean(log(rel_density), na.rm=TRUE))) / sd(log(rel_density), na.rm=TRUE)) %>%
    ungroup()
  
  dat$datePosix <- as.POSIXlt(strptime(dat$date, "%m/%d/%Y %H:%M:%S"))
  dat$doy <- lubridate::yday(dat$datePosix)
  
  speciesList <- unique(dat$species)
  
  save(dat, sinc, file =paste0("D:/projects/ecoDrought/bodySize/All_Size_inputs/Size_inputs_", date,"/dat.R"))
  
  rm(all_fish_init, all_fish, dat0)
}
```

```{r load data if no new data}
if(!newData){
  load(paste0("D:/projects/ecoDrought/bodySize/All_Size_inputs/Size_inputs_", date,"/dat.R"))
  speciesList <- unique(dat$species)
}
```

```{r theme_publication}

theme_publication <- function(base_size=14, base_family="") {
  library(grid)
  library(ggthemes)
  (theme_foundation(base_size=base_size, base_family=base_family)
    + theme(plot.title = element_text(face = "bold",
                                      size = rel(1.2), hjust = 0.5),
            text = element_text(),
            panel.background = element_rect(colour = NA),
            plot.background = element_rect(colour = NA),
            axis.title = element_text(face = "bold",size = rel(1)),
            axis.title.y = element_text(angle=90,vjust =2),
            axis.title.x = element_text(vjust = -0.2),
            axis.text = element_text(),
            axis.line = element_line(colour="black"),
            axis.ticks = element_line(),
            panel.grid.major = element_line(colour="#f0f0f0"),
            panel.grid.minor = element_blank(),
            legend.key = element_rect(colour = NA),
            legend.position = "bottom",
            legend.direction = "horizontal",
            legend.key.size= unit(0.2, "cm"),
            #    legend.margin = unit(0, "cm"),
            legend.title = element_text(face="italic"),
            plot.margin=unit(c(10,5,5,5),"mm"),
            strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
            strip.text = element_text(face="bold"),
            panel.border = element_rect(colour = "black")
    ))

}
```


***

## 1. Spatial distribution of length by species
Spatially-nested lmer() to explore spatial structuring of variation in body sizes
HUC12 nested within HUC8 nested within Location

***
```{r joy plots}
ggplot(dat %>% filter(length < 700), aes(x = length, y = species)) +
  geom_density_ridges() +
  theme_publication()


```



```{r spatially-nested lmer by species}

# build up hierarchical random effects for each species 
## Hierarchical graphs for huc8/huc12/location 
  lmerModel <- function(df) {
    lmer(log10Length ~ 1 + (1|location/HUC8_Name/HUC12_NAME), data = df) 
  }

  reBySpecies <- function(mod, species){
    re <- as.data.frame(ranef(mod))

    re12 <- re %>%
      filter(grpvar == 'HUC12_NAME:(HUC8_Name:location)') %>%
      mutate(huc12 = str_split(grp, ":", simplify = TRUE)[,1],
             huc8 = str_split(grp, ":", simplify = TRUE)[,2],
             loc = str_split(grp, ":", simplify = TRUE)[,3]) %>%
      rename(condval12 = condval)
  
    re8 <- re %>%
      filter(grpvar == 'HUC8_Name:location') %>%
      mutate(loc = str_split(grp, ":", simplify = TRUE)[,2],
             huc8 = str_split(grp, ":", simplify = TRUE)[,1]) %>%
      rename(condval8 = condval)
    
    reLoc <- re %>%
      filter(grpvar == 'location') %>%
      mutate(loc = grp) %>%
      rename(condvalLoc = condval)%>% 
      select(condvalLoc, loc)
    
    reAll1 <- left_join(re8, reLoc) %>% select(condval8, loc, huc8, condvalLoc)
    reAll <- left_join(re12, reAll1) %>% arrange(loc, huc8, huc12)
    reAll <- reAll %>%
      mutate(condvalLoc8 = condvalLoc + condval8,
             condvalLoc812 = condvalLoc + condval8 + condval12)
    
    return(reAll)
  }
  
  hierRECharts <- function(d, species) {
      gg <- ggplot(d, aes(condvalLoc812, huc12)) +
      geom_point(color = '#ff9a36') +
      geom_point(aes(condvalLoc8, huc12), color = '#3f7bc4') +
      geom_point(aes(condvalLoc, huc12, shape = factor(loc)), color = '#bdb2a8') +
      geom_vline(xintercept = 0) +
      scale_y_discrete(guide = guide_none(title = NULL)) +
      scale_x_continuous("Length log10") +
      ggtitle(species) +
      facet_wrap(~huc8)
      
    ggsave(paste0("D:/projects/ecoDrought/bodySize/All_Size_inputs/ecoDrought_SizeAnalysis/charts/hierResids_",species,".png"))
    return(gg)
  }
  
##  
    # Charts of nested random effects 
  datBySpeciesRE <- dat %>%
    filter(species %in% speciesList[c(1,3,4,5,7)]) %>%
    #filter(species %in% speciesList[5]) %>% #
    group_by(species) %>% 
    nest() %>%
    mutate(model = map(data, lmerModel)) %>%
    mutate(RE = map2(model, species, reBySpecies)) %>%
    mutate(charts = map2(RE, species, hierRECharts))
    
  datBySpeciesRE$charts
  
  ## SDs of nested levels ################
  datBySpeciesRESD <- datBySpeciesRE %>%
    mutate(tidy = map(model, broom::tidy)) %>% 
    dt_unnest(tidy) %>%
    select(species, estimate, term, group) %>%
    filter(!term == '(Intercept)')
  
  ggSD <- ggplot(datBySpeciesRESD, aes(species,estimate, fill = group)) +
    geom_col(position = "dodge") +
    scale_y_continuous("Standard deviation of nested lengths")
  ggSD
  
  ## plot RE across species
  datBySpeciesRE_RE <- datBySpeciesRE %>%
    dt_unnest(RE)

  # by loc
  ggplot(datBySpeciesRE_RE, aes(condvalLoc812, species)) +
    geom_point(color = '#ff9a36', alpha = 0.2, position = position_nudge(y = 0.2)) +
    geom_point(aes(condvalLoc8, species), color = '#3f7bc4', alpha = 0.2, position = position_nudge(y = -0.2)) +
    geom_point(aes(condvalLoc, species), color = 'black') +
    geom_vline(xintercept = 0) +
    scale_x_continuous("Length log10") +
    facet_wrap(~loc)

  # by huc8
  nByhuc8 <- datBySpeciesRE_RE %>%
    group_by(huc8) %>%
    summarise(nSpp = length(unique(species)))
  
  datBySpeciesRE_RE_SppN <- left_join(datBySpeciesRE_RE, nByhuc8)
  
  ggplot(datBySpeciesRE_RE_SppN %>% filter(nSpp >1), aes(condvalLoc812, species)) +
    geom_point(color = '#ff9a36', alpha = 0.2, position = position_nudge(y = 0.2)) +
    geom_point(aes(condvalLoc8, species), color = '#3f7bc4', position = position_nudge(y = -0.2)) +
    geom_point(aes(condvalLoc, species), color = 'black') +
    geom_vline(xintercept = 0) +
    scale_x_continuous("Length log10") +
    facet_wrap(loc ~ huc8, labeller = labeller(.multi_line = FALSE))
  
##############################################################################  
  # some species have data for one location only #2 6 8 9
  # Colorado River Cutthroat Trout model has singularity - ignoring for now
  ## Hierarchical graphs for huc12/huc8
  
  lmerModelNoLocation <- function(df) {
    lmer(log10Length ~ 1 + (1|HUC8_Name/HUC12_NAME), data = df)
  }
  
  # build up hierarchical random effects for each species 
  reBySpecies812 <- function(mod, species){
    re <- as.data.frame(ranef(mod))

    re12 <- re %>%
      filter(grpvar == 'HUC12_NAME:HUC8_Name') %>%
      mutate(huc12 = str_split(grp, ":", simplify = TRUE)[,1],
             huc8 = str_split(grp, ":", simplify = TRUE)[,2]
             ) %>%
      rename(condval12 = condval)

    re8 <- re %>%
      filter(grpvar == 'HUC8_Name') %>%
      mutate(huc8 = grp) %>%
      rename(condval8 = condval) %>% 
      select(condval8, huc8)
    
    reAll <- left_join(re12, re8) %>% arrange(huc8, huc12)
    reAll <- reAll %>%
      mutate(condval8 = condval8,
             condval812 = condval8 + condval12)
    
    return(reAll)
  }
  
  hierREChartsNoLocation <- function(d, species) {
    gg <- ggplot(d, aes(condval812, huc12)) +
      geom_point(color = '#ff9a36') +
      geom_point(aes(condval8, huc12), color = '#3f7bc4') +
      geom_vline(xintercept = 0) +
      scale_y_discrete(guide = guide_none(title = NULL)) +
      scale_x_continuous("Length") +
      ggtitle(species) +
      facet_wrap(~huc8)
    
    ggsave(paste0("D:/projects/ecoDrought/bodySize/All_Size_inputs/ecoDrought_SizeAnalysis/charts/hierResids_",species,".png"))
    
    return(gg)
  }  
  
  # Charts of nested random effects 
  datBySpeciesRENoLocation <- dat %>%
    filter(species %in% speciesList[c(2,8,9)]) %>%
    group_by(species) %>% 
    nest() %>%
    mutate(model = map(data, lmerModelNoLocation)) %>%
    mutate(RE = map2(model, species, reBySpecies812)) %>%
    mutate(charts = map2(RE, species, hierREChartsNoLocation))
    
  datBySpeciesRENoLocation$charts 
  
  ## SDs of nested levels ################
  datBySpeciesRENoLocationSD <- datBySpeciesRENoLocation %>%
    mutate(tidy = map(model, broom::tidy)) %>% 
    dt_unnest(tidy) %>%
    select(species, estimate, term, group) %>%
    filter(!term == '(Intercept)')
  
  ggNoLocationSD <- ggplot(datBySpeciesRENoLocationSD, aes(species,estimate, fill = group)) +
    geom_col(position = "dodge") +
    scale_y_continuous("Standard deviation of length") 
  ggNoLocationSD 
  
  ## plot RE across species
  datBySpeciesRENoLocation_RE <- datBySpeciesRENoLocation %>%
    dt_unnest(RE)

  # by huc8
  nNoLocationByhuc8 <- datBySpeciesRENoLocation_RE %>%
    group_by(huc8) %>%
    summarise(nSpp = length(unique(species)))
  
  datBySpeciesRENoLocation_RE_SppN <- left_join(datBySpeciesRENoLocation_RE, nNoLocationByhuc8)
  
  ggplot(datBySpeciesRENoLocation_RE_SppN %>% filter(nSpp >1), aes(condval812, species)) +
    geom_point(color = '#ff9a36', alpha = 0.2, position = position_nudge(y = 0.2)) +
    geom_point(aes(condval8, species), color = '#3f7bc4', position = position_nudge(y = -0.2)) +
    geom_vline(xintercept = 0) +
    scale_x_continuous("Length") +
    facet_wrap( ~ huc8)

```



